Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2022-03-07 paper/paper.tex
WIP on laptop

An assumption made in these papers is that all EphA receptor subtypes
contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which
increases from a low level at the nasal side of the retina, to a high level at
the temporal side. The precise contribution of each subtype to the overall
level is assumed not to be important. In our model, receptor $r_0$ increases
in the retina in a nasal to temporal direction (Fig\,\ref{f:expressions}B) and
so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et
al.~experiments. In half of the RGCs, we knock-in EphA3 by adding to the value
of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote
EphA3 ki/ki). To add the additional knock-down of EphA4
in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all}
RGCs.

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
\label{f:GJreber}
\end{figure}

@@

One problem with axon gradient-following is the ability of the axon to
successfully detect the direction of the gradient given the small distances
over which the axon can sample signalling molecules and the size of receptor
binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
system to self-organise in the presence of noise, by introducing a stochastic
element to each branch's sampling of the gradient.

@@

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations_GClikeGJ.png}
\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
optimization, genetic manipulations. As for the surgical manipulations, the
results are similar to those of the $\mathbf{GJ}$ model with the equivalent
parameters. However, there are subtle differences such as XXX.}
\label{f:GClikeGJreber}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
\label{f:GJeph}
\end{figure}

An assumption made in these papers is that all EphA receptor subtypes
contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which
increases from a low level at the nasal side of the retina, to a high level at
the temporal side. The precise contribution of each subtype to the overall
level is assumed not to be important. In our model, receptor $r_0$ increases
in the retina in a nasal to temporal direction (Fig\,\ref{f:expressions}B) and
so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et
al.~experiments. In half of the RGCs, we knock-in EphA3 by adding to the value
of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote
EphA3 ki/ki). To add the additional knock-down of EphA4
in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all}
RGCs. Fig.\,\ref{f:GJeph}A-C shows, for the single EphA3 knock-in (EphA3
ki/+), branch positions, axon centroids and final rostral-caudal position on
the tectum plotted versus the originating nasal-temporal retinal
coordinate. In A, branches are coloured in the red/green duochrome scheme
according to their origin location on the retina. In B \& C, axon centroids of
RGCs which were treated with the knock-in manipulation are coloured red,
wildtype cells are blue. The map of treated cells on the tectum is shifted
slightly rostrally, but the shift becomes statistically insignificant at about
0.75 along the nasal-temporal axis. While the map collapse is less abrupt than
in the experimental result of \citet{brown_topographic_2000}, it is comparable
with the result from \citet{simpson_simple_2011}.  Fig.\,\ref{f:GJeph}D-F
shows the same trio of graphs for the homozygote double knock-in, EphA3
ki/ki. Here, the treated axons are pushed further towards the rostral end of
the tectum and the untreated axons are pushed caudally. The maps are
completely separated; as in the experiment, there is no `collapse point'.

The rostral movement of treated axons occurs due to their increased expression
of EphA ($r_0$), which increases the interaction with the tectal
rostral-caudal ligand gradient. The opposing receptor expression ($r_2$) is
unchanged by the manipulation and so those axons with increased $r_0$
expression are pulled to a more rostral position than they would otherwise
take. The caudal movement of wildtype axons, whose receptor expression in
Figs.\,\ref{f:GJeph}A-F is unchanged, occurs due to competition with the
caudally-drawn treated axons.
%
%Interestingly, the feature of a slight 'elbow' in the double knockin
%graph (F) at about 0.7 on the nasal-temporal axis matches that seen in the
%experimental result. Well, kinda.

Figs.\,\ref{f:GJeph}G-I show the results for the Reber et al.~manipulation, in
which the EphA expression is selectively knocked in in 50\% of cells and
knocked down in all cells. The collapse point has moved from about 0.75 in
Fig.\,\ref{f:GJeph}C to about 0.85 (\textbf{check}), in agreement with the
experimental result. However, the significant rostral movement of the treated
axons seen in \citet{reber_relative_2004}, where not a single treated axon was
found more than 30\% along the rostral-caudal axis, was absent from our result.

@@

One problem with axon gradient-following is the ability of the axon to
successfully detect the direction of the gradient given the small distances
over which the axon can sample signalling molecules and the size of receptor
binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
chemoaffinity-only and the $\mathbf{GJ}$ model systems to self-organise in the
presence of noise, by introducing a stochastic element to each branch's
sampling of the gradient.

@@

% What about the EphA manipulations with noise? Ans: No strong effect (just
% add "noise_gain" : 0.5, to one of the expt jsons.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations_GClikeGJ.png}
\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
optimization, genetic manipulations. As for the surgical manipulations, the
results are similar to those of the $\mathbf{GJ}$ model with the equivalent
parameters. However, there are subtle differences such as XXX.}
\label{f:GClikeGJeph}
\end{figure}
